A comprehensive framework to capture the arcana of neuroimaging analysis

### 2018-10-17 manuscript/manuscript.tex
added in references and fixed vein images so that they show up in correct aspect

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{natbib}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{color}

@@

% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}

@@

%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{A comprehensive framework to capture the arcana of neuroimaging
analysis%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}

@@

\authorrunning{
        Thomas G. Close \and
	Phillip G. D. Ward  \and
	Francesco Sforazzini  \and
	Wojtek Goscinski  \and
	Zhaolin Chen  \and
	Gary F. Egan  \and} % if too long for running head

@@

Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.

\keywords{Neuroimaging \and Workflows \and Repository \and Large-scale \and Reproducibility}

% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}

@@

A barrier to designing portable and complete workflows is the
heterogeneity of data storage conventions \citep{marcus_extensible_2007,das_loris:_2012,gorgolewski_brain_2016}.
To address this, the emerging Brain
Imaging Data Standard (BIDS) \citep{gorgolewski_brain_2016} provides a way
to standardise the storage of neuroimaging data in file system
directories. BIDS specifies strict file and directory naming
conventions, which facilitate the design of portable \emph{BIDS Apps}
(https://bids-apps.neuroimaging.io). However, for research groups with
sufficient informatics support, software-managed repositories \citep{marcus_extensible_2007,das_loris:_2012} can provide additional features, such as
flexible access-control and automated pipelines. Therefore, for
published workflows, the choice of repository should be transparent in
order to maximise their audience.

@@

Our objective was to extract common elements of repository-centric
workflow design into an abstract framework to make it practical to
implement flexible, portable and complete workflows for a wide range of
neuroimaging analyses. \emph{Abstraction of Repository-Centric ANAlysis
(Arcana)} \\(\href{http://arcana.readthedocs.io}{http://arcana.readthedocs.io} is a Python framework for
designing complex workflows in which modular Nipype pipelines operate on
data stored in various repositories. Intermediate derivatives are
derived on demand, checking against stored provenance for required
updates. Analyses can be applied to XNAT, BIDS and plain-directory
repositories, and using Nipype's execution plugins, run on workstations
or be submitted as batch jobs to HPC schedulers. Arcana's architecture,
with programmatic workflow construction---yet clear delineation between
analysis design and application---facilitates the implementation of
complex workflows that are portable and complete.

@@

The \emph{ModulesEnvironment} module can be used if environment
modules \citep{furlani_modules:_1991} are installed on the system (typical in many
HPC environments). In this case, modules are loaded before the interface of a pipeline node
is run based on the requirements specified for the node during construction of the pipeline
(see \emph{Pipeline constructors} and then unloaded afterwards. Mappings from non-standard
module names installed on the system to requirements required by the Study can be
can be passed to the \emph{packages\_map} keyword argument of the ModulesEnvironment
at initialisation. Likewise non-standard versions for packages can be mapped onto requirements
using the \emph{versions\_map} keyword argument.

@@

The Arcana framework is distributed as a publicly available software
package via GitHub (\url{github.com/MonashBI/arcana}) and the
Python Package Index (PyPI) (\url{pypi.org/project/arcana/}).
Study classes for T1, T2* and diffusion weighted MRI data have been
implemented as part of the \emph{Biomedical imAgiNg ANAlysis (Banana)} package
(\url{github.com/MonashBI/banana}; \url{pypi.org./project/banana}).
All three classes, \emph{T1Study}, \emph{T2starStudy} and
\emph{DmriStudy}, inherit generic image analysis methods, such as
registration and brain extraction, from the base class \emph{MRIStudy}.
The T1Study and T2starStudy classes are aggregated into the
\emph{T2starwT1Study} MultiStudy class to combine analysis of T2* and
T1-weighted contrasts.

@@

The \emph{T2starwT1Study} class implements an algorithm to generate
\emph{composite vein images} from the combination of QSM and SWI
contrasts derived from the T2*-weighted acquisition with an vein atlas
registered to the T1-weighted acquisition \citep{ward_combining_2018} (Figure \ref{fig:t2star_t1_study}).

@@

To test the analysis workflows implemented in ArcanaPaper, a healthy
volunteer was scanned using a 3T Siemens Skyra with a 32-channel head
and neck coil. The protocol was a T1-weighted MPRAGE (1mm contiguous,
matrix size 256x240x192, FOV 256x240x192, TE = 2.13ms, TR, = 2300ms, TI
= 900ms, bandwidth=230Hz/pixel), GRE (1.8 mm contiguous, matrix size
256x232x72, FOV 230x208x130, TE = 20ms, TR, = 30ms,
bandwidth=120Hz/pixel), diffusion MRI (1.2 mm contiguous, matrix size
110x100x60, FOV 256x240x192, TE = 95ms, TR, = 8200ms, 33 diffusion
directions with b=1500 mm2/s and 3 b=0, bandwidth=781Hz/pixel)
(\textbf{cite Lifespan paper ??}). Methods for Figure \ref{fig:veins}, \ref{fig:fa_acs},
and \ref{fig:tractography} in
the ArcanaPaper class were applied to the acquired dataset to using the
\emph{SlurmProcessor} to submit jobs to the the MASSIVE HPC cluster
\citep{goscinski_multi-modal_2014} in order to generate the figures.

Figure \ref{fig:veins} displays composite vein images and vein masks for the healthy
volunteer along with the SWI and QSM intermediate derivatives and the
input vein atlas. The derived vein images are comparable to those
generated by the original implementation \citep{ward_combining_2018}. The complete
workflow took ?? min to complete over ?? compute cores.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/veins}
  \caption{Vein image (\emph{right column}) constructed by
combining susceptibility weighted imaging (SWI) (\emph{left column}),
quantitative susceptibility mapping (QSM) (\emph{left-middle column})
and a vein atlas from manual tracings (\emph{right-middle column}).
\emph{Top row}: sagittal slices. \emph{Middle row}: axial slices.
\emph{Bottom row}: coronal slices.}
\label{fig:veins}
\end{figure*}

@@

The flexibility and portability of the Arcana framework increases the
feasibility of community collaborations on workflow implementations. The
improvement of code quality in larger community efforts, due to more
eyeballs to detect and fix errors, has the potential to form a
reinforcing cycle where more developers are attracted to the project.
To these ends, the Banana code repository on GitHub
(\url{http://github.com/MonashBI/banana.git}) is proposed as a code base
for communal development of biomedical imaging workflows using Arcana.

@@

Ensuring that the consistent versions of external tools are used
throughout the analysis is important to avoid introducing biases due to
algorithm updates. In systems with environment modules \citep{furlani_modules:_1991}
installed, Arcana can load and unload the required modules before and
after each node is executed. When running Arcana within a container,
environment modules and software versions can be installed inside the
container giving exact control over the versions used. To these ends, a
Docker container is available on Docker Hub,
(\href{http://hub.docker.com/r/monashbi/banana}{{http://hub.docker.com/r/monashbi/banana}}),
which can be used as a base for biomedical imaging analysis containers.

While the same tools and versions should be applied across an analysis
to avoid bias, there are cases where it is desirable to rerun the same
analysis with different tools substituted at various points in the
workflow. In particular, when introducing new tools or upgrades to
existing tools, it is important to show the effect on the final results
in comparison with existing methods. Furthermore, it is typically not
clear what variability between results produced by comparable tools is
due to. Therefore, in the absence of \emph{a priori} reason to favour a
particular tool, perhaps the most rigorous approach is to rerun analyses
with different combinations of available tools and only present results
that are robust to the ``analytic noise'' (cite here?) they introduce.
Switch parameters make it straightforward to rerun analyses in Arcana
with substituted tools while controlling all other aspects of the
workflow.

@@

\begin{acknowledgements}
The authors acknowledge the facilities and scientific and technical assistance
of the National Imaging Facility, a National Collaborative Research
Infrastructure Strategy (NCRIS) capability, at Monash Biomedical Imaging,
Monash University. The ``transparent repository'' feature of Arcana was
inspired by in-house software written by Parnesh Raniga while he was
employed at Monash Biomedical Imaging prior to 2016.
\end{acknowledgements}

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required

\RequirePackage{fix-cm}

%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn

\smartqed  % flush right qed marks, e.g. at end of proof

% \usepackage{mathptmx}      % use Times fonts if available on your TeX system

\usepackage{natbib}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{color}

@@

@@

% Insert the name of "your journal" with
% \journalname{myjournal}

\begin{document}

\title{A comprehensive framework to capture the arcana of neuroimaging
analysis
}

@@

\authorrunning{
     Thomas G. Close \and
	Phillip G. D. Ward  \and
	Francesco Sforazzini  \and
	Wojtek Goscinski  \and
	Zhaolin Chen  \and
	Gary F. Egan}

@@

These are the notes I have for my abstract

\begin{itemize}
\item The heterogeneous collection of neuroimaging analysis tools, parameters, styles makes it difficult to design comprehensive workflows
\item Complex workflows in a flexible extensible manner, reusable
\item portable between different data storage solutions/scale of studies, separation between design and application yet programmatic
\item Easily updatable, switchable between tools/versions
\end{itemize}

\keywords{Neuroimaging \and Workflows \and Repository \and Reproducibility \and Reusability \and Large-scale \and Python}

@@

A barrier to designing portable and complete workflows is the
heterogeneity of data storage conventions \citep{marcus_extensible_2007,das_loris:_2012,gorgolewski_brain_2016}.
To address this, the emerging Brain
Imaging Data Standard (BIDS) \citep{gorgolewski_brain_2016} provides a way
to standardise the storage of neuroimaging data in file system
directories. BIDS specifies strict file and directory naming
conventions, which facilitate the design of portable \emph{BIDS Apps}
(\url{bids-apps.neuroimaging.io}). However, for research groups with
sufficient informatics support, software-managed repositories \citep{marcus_extensible_2007,das_loris:_2012}
can provide additional features, such as
flexible access-control and automated pipelines. Therefore, for
published workflows, the choice of repository should be transparent in
order to maximise their audience.

@@

Our objective was to extract common elements of repository-centric
workflow design into an abstract framework to make it practical to
implement flexible, portable and complete workflows for a wide range of
neuroimaging analyses. \emph{Abstraction of Repository-Centric ANAlysis
(Arcana)} (\url{arcana.readthedocs.io}) is a Python framework for
designing complex workflows in which modular Nipype pipelines operate on
data stored in various repositories. Intermediate derivatives are
derived on demand, checking against stored provenance for required
updates. Analyses can be applied to XNAT, BIDS and plain-directory
repositories, and using Nipype's execution plugins, run on workstations
or be submitted as batch jobs to HPC schedulers. Arcana's architecture,
with programmatic workflow construction---yet clear delineation between
analysis design and application---facilitates the implementation of
complex workflows that are portable and complete.

@@

The \emph{ModulesEnvironment} module can be used if environment
modules \citep{furlani_modules:_1991} are installed on the system (typical in many
HPC environments). In this case, modules are loaded before a workflow node is run,
based on the requirements specified for the node during construction of the pipeline
(see \emph{Pipeline constructors}), and then unloaded afterwards. Mappings from non-standard
module names installed on the system to requirements required by the Study can be
can be passed to the \emph{packages\_map} keyword argument of the ModulesEnvironment
at initialisation. Likewise non-standard versions for packages can be mapped onto requirements
using the \emph{versions\_map} keyword argument.

@@

The Arcana framework is distributed as a publicly available software
package via GitHub (\url{github.com/MonashBI/arcana}) and the
Python Package Index (PyPI) (\url{pypi.org/project/arcana/}).
Study classes for T1, T2* and diffusion weighted MRI data have been
implemented as part of the \emph{Biomedical imAgiNg ANAlysis (Banana)} package
(\url{github.com/MonashBI/banana}; \url{pypi.org./project/banana}).
All three classes, \emph{T1Study}, \emph{T2starStudy} and
\emph{DmriStudy}, inherit generic image analysis methods, such as
registration and brain extraction, from the base class \emph{MriStudy}.
The T1Study and T2starStudy classes are aggregated into the
\emph{T2starT1Study} MultiStudy class to combine analysis of T2* and
T1-weighted contrasts.

@@

The \emph{T2starT1Study} class implements an algorithm to generate
\emph{composite vein images} from the combination of QSM and SWI
contrasts derived from the T2*-weighted acquisition with an vein atlas
registered to the T1-weighted acquisition \citep{ward_combining_2018}.
(Figure \ref{fig:t2star_t1_study}).

@@

To test the analysis workflows implemented in ArcanaPaper, a healthy
volunteer was scanned using a 3T Siemens Skyra with a 32-channel head
and neck coil. The protocol was a T1-weighted MPRAGE (1mm contiguous,
matrix size 256x240x192, FOV 256x240x192, TE = 2.13ms, TR, = 2300ms, TI
= 900ms, bandwidth=230Hz/pixel), GRE (1.8 mm contiguous, matrix size
256x232x72, FOV 230x208x130, TE = 20ms, TR, = 30ms,
bandwidth=120Hz/pixel), diffusion MRI (1.2 mm contiguous, matrix size
110x100x60, FOV 256x240x192, TE = 95ms, TR, = 8200ms, 33 diffusion
directions with b=1500 mm2/s and 3 b=0, bandwidth=781Hz/pixel).
Methods for Figure \ref{fig:veins}, \ref{fig:fa_adc}, and \ref{fig:tractography} in
the ArcanaPaper class were applied to the acquired dataset to using the
\emph{SlurmProcessor} to submit jobs to the the MASSIVE HPC cluster
\citep{goscinski_multi-modal_2014} in order to generate the figures.

Figure \ref{fig:veins} displays composite vein images for the healthy
volunteer along with the SWI and QSM intermediate derivatives and the
input vein atlas. The derived vein images are comparable to those
generated by the original implementation \citep{ward_combining_2018}.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/veins}
  \caption{Composite vein image (\emph{third row}) constructed by
  combining susceptibility weighted imaging (SWI) (\emph{top row}),
  quantitative susceptibility mapping (QSM) (\emph{second row})
  and a vein atlas from manual tracings. A vein mask was then
  generated (\emph{bottom row}) by thresholding the composite
  vein image.
  \emph{left column}: axial slices. \emph{centre column}: coronal slices.
  \emph{right column}: sagittal slices.}
\label{fig:veins}
\end{figure*}

@@

The flexibility and portability of the Arcana framework increases the
feasibility of community collaborations on workflow implementations. The
improvement of code quality in larger community efforts, due to more
eyeballs to detect and fix errors, has the potential to form a
reinforcing cycle where more developers are attracted to the project.
To these ends, the Banana code repository on GitHub
(\url{github.com/MonashBI/banana.git}) is proposed as a code base
for communal development of biomedical imaging workflows using Arcana.

@@

Ensuring that the consistent versions of external tools are used
throughout the analysis is important to avoid introducing biases due to
algorithm updates. In systems with environment modules \citep{furlani_modules:_1991}
installed, Arcana can load and unload the required modules before and
after each node is executed. When running Arcana within a container,
environment modules and software versions can be installed inside the
container giving exact control over the versions used. To these ends, a
Docker container is available on Docker Hub,
(\url{hub.docker.com/r/monashbi/banana}), which can be used as a
base for biomedical imaging analysis containers. In future versions of Arcana,
additional Environment modules could be implemented to run each pipeline
node within its own container if required to take advantage of
containers maintained by tool developers (e.g. 
\url{hub.docker.com/r/vistalab/freesurfer/}).

While the same tools and versions should be applied across an analysis
to avoid bias, there are cases where it is desirable to rerun the same
analysis with different tools substituted at various points in the
workflow. In particular, when introducing new tools or upgrades to
existing tools, it is important to show the effect on the final results
in comparison with existing methods. Furthermore, it is typically not
clear what variability between results produced by comparable tools is
due to. Therefore, in the absence of \emph{a priori} reason to favour a
particular tool, perhaps the most rigorous approach is to rerun analyses
with different combinations of available tools and only present results
that are robust to the ``analytic noise'' \citep{maumet:inserm-01886089} they introduce.
Switch parameters make it straightforward to rerun analyses in Arcana
with substituted tools while controlling all other aspects of the
workflow.

@@

\begin{acknowledgements}
The authors acknowledge the facilities and scientific and technical assistance
of the National Imaging Facility, a National Collaborative Research
Infrastructure Strategy (NCRIS) capability, at Monash Biomedical Imaging,
Monash University. The ``transparent repository'' feature of Arcana was
inspired by in-house software written by Parnesh Raniga while he was
employed at Monash University prior to 2016.
\end{acknowledgements}
